---
title: "visualization_2"
output: github_document
date: "2025-09-30"
---

```{r}
library(tidyverse)
library(patchwork)
library(p8105.datasets)
```


import the weather data

```{r}
data("weather_df")
```

let's make our basic scatterplot
note: `labs` function to add descriptions to your plot
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y=tmax)) +
  geom_point(aes(color = name), alpha = 0.5) +
  labs(
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    title = "Temp scatterplot",
    caption = "Data from NOAA",
    color = "Location"
  )
```

## Scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y=tmax)) +
  geom_point(aes(color = name), alpha = 0.5) +
  labs(
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    title = "Temp scatterplot",
    caption = "Data from NOAA",
    color = "Location"
  ) +
  scale_x_continuous(
    breaks = c(-20, 0, 25),
    labels = c("-20C", "0", "25")
  ) +
  scale_y_continuous(
    trans = "sqrt",
    limits = c(10,30)
  )
```

another way to zoom in to an area of your plot...
```{r}
weather_df %>%
  filter(tmax > 10, tmax <30) %>% 
  ggplot(aes(x = tmin, y=tmax)) +
  geom_point(aes(color = name), alpha = 0.5) +
  labs(
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    title = "Temp scatterplot",
    caption = "Data from NOAA",
    color = "Location"
  ) +
  scale_x_continuous(
    breaks = c(-20, 0, 25),
    labels = c("-20C", "0", "25")
  )  
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y=tmax)) +
  geom_point(aes(color = name), alpha = 0.5) +
  labs(
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    title = "Temp scatterplot",
    caption = "Data from NOAA",
    color = "Location"
  ) +
  scale_x_continuous(
    breaks = c(-20, 0, 25),
    labels = c("-20C", "0", "25")
  ) +
  viridis::scale_color_viridis(
    discrete = TRUE
  )
```


## Themes

make my base plot.
```{r}
ggplot_temp = 
  weather_df %>% 
  ggplot(aes(x = tmin, y=tmax)) +
  geom_point(aes(color = name), alpha = 0.5) +
  labs(
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    title = "Temp scatterplot",
    caption = "Data from NOAA",
    color = "Location"
  ) +
  scale_x_continuous(
    breaks = c(-20, 0, 25),
    labels = c("-20C", "0", "25")
  ) +
  viridis::scale_color_viridis(
    discrete = TRUE
  )
```

update my base plot
note: order of code matters here
```{r}
ggplot_temp +
  theme_dark() +
  theme(legend.position = "bottom") 

ggplot_temp +
  theme_minimal() +
  theme(legend.position = "bottom") 

ggplot_temp +
  theme_bw() +
  theme(legend.position = "bottom") 
```

```{r}
ggsave("weather_scatterplot.png", ggplot_temp)
```


## Adding data in geoms

overlapping two different data sets
```{r}
central_park_df = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

molokai_df =
 weather_df %>% 
  filter(name == "Molokai_HI")

ggplot(data = molokai_df, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park_df)
```

## `pathwork`

making three plots and combining them (aka putting them side by side) using `patchwork`
```{r}
ggplot_tmax_tmin =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  theme(legend.position = "none")

ggplot_prec_density =
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5) +
  theme(legend.position = "none")

ggplot_temp_season =
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom")

(ggplot_tmax_tmin + ggplot_prec_density) / ggplot_temp_season
```


## Data manipulation

let's make temp. violin plots.
note: using `mutate(name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole, WA")))` I'm putting Molokai_HI 1st, CentralPark_NY 2nd, and Waterhole, WA 3rd 
```{r}
weather_df %>% 
  mutate(name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole, WA"))) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.5)
```

another way to do this
```{r}
weather_df %>% 
  mutate(name = fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.5)
```

what about data tidiness??

note: `mutate(vist = fct_inorder(visit))` = what order does it appear in the first dataset we started with
```{r}
pulse_df = 
  haven::read_sas("data/public_pulse_data.sas7bdat") %>% 
  janitor:: clean_names() %>% 
  pivot_longer(
    cols = bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) %>% 
  mutate(vist = fct_inorder(visit))

pulse_df %>% 
  ggplot(aes(x = visit, y = bdi)) +
  geom_boxplot()
```


Make a plot for the FAS study.

```{r}
pups_df = 
  read_csv("data/FAS_pups.csv", na = c("NA", ".", ""), skip = 3) %>% 
  janitor::clean_names() %>% 
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female"
    )
  )

litters_df =
  read_csv("data/FAS_litters.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>%  
  separate(group, into = c("dose", "tx_day", sep = 3))

fas_df = 
  left_join(pups_df, litters_df, by = "litter_number")

fas_df %>% 
  select(pd_ears:tx_day) %>% 
  pivot_longer(
    cols = pd_ears:pd_walk,
    names_to = "outcome",
    names_prefix = "pd_",
    values_to = "pn_day") %>% 
  mutate(outocme = fct_reorder(outcome, pn_day)) %>%
  ggplot(aes(x = dose, y = pn_day)) +
  geom_violin() +
  facet_grid(tx_day ~ outcome)
```

